---
title: "SVI in Livingston County vs. NYS"
author: "Daniel Noone"
editor: source
format: dashboard
---
# Theme Overview

```{r}
library(tidyverse)
library(ggthemes)
library(reshape2)
svi <-  read.csv("C:\\Users\\noone\\OneDrive\\Documents\\School\\SUNY Geneseo\\Classes\\Spring 2025\\DANL499\\SVI2022_NEWYORK_tract.csv")

svi_clean <- svi %>%
  mutate(RPL_THEME1 = ifelse(RPL_THEME1 == -999, NA, RPL_THEME1),
         RPL_THEME2 = ifelse(RPL_THEME2 == -999, NA, RPL_THEME2),
         RPL_THEME3 = ifelse(RPL_THEME3 == -999, NA, RPL_THEME3),
         RPL_THEME4 = ifelse(RPL_THEME4 == -999, NA, RPL_THEME4),
         RPL_THEMES = ifelse(RPL_THEMES == -999, NA, RPL_THEMES))


liv_avg <- svi_clean %>%
  filter(COUNTY == "Livingston County") %>%
  summarize(
    livingston_RPL_THEME1 = mean(RPL_THEME1, na.rm = TRUE),
    livingston_RPL_THEME2 = mean(RPL_THEME2, na.rm = TRUE),
    livingston_RPL_THEME3 = mean(RPL_THEME3, na.rm = TRUE),
    livingston_RPL_THEME4 = mean(RPL_THEME4, na.rm = TRUE),
    livingston_RPL_THEMES = mean(RPL_THEMES, na.rm = TRUE)
  )


nys_avg <- svi_clean %>%
  summarize(
    nys_RPL_THEME1 = mean(RPL_THEME1, na.rm = TRUE),
    nys_RPL_THEME2 = mean(RPL_THEME2, na.rm = TRUE),
    nys_RPL_THEME3 = mean(RPL_THEME3, na.rm = TRUE),
    nys_RPL_THEME4 = mean(RPL_THEME4, na.rm = TRUE),
    nys_RPL_THEMES = mean(RPL_THEMES, na.rm = TRUE)
  )


comp_df <- bind_cols(liv_avg, nys_avg)

summary <- svi_clean %>%
  summarise(
    T1 = mean(RPL_THEME1, na.rm = TRUE),
    T2 = mean(RPL_THEME2, na.rm = TRUE),
    T3 = mean(RPL_THEME3, na.rm = TRUE),
    T4 = mean(RPL_THEME4, na.rm = TRUE),
    COMP = mean(RPL_THEMES, na.rm = TRUE)
  )

```


```{r}
#| echo: false
#| message: false
#| warning: false
library(knitr)
kable(summary, caption = "Livingston County Theme Averages")
```


```{r}
#| content: valuebox
#| title: "Theme 1 (Socioeconomic Status)"

list(
  icon = "money-bill",
  color = "success",
  value = paste0(round(comp_df$livingston_RPL_THEME1,2))
)

```

```{r}
#| content: valuebox
#| title: "Theme 2 (Household Characteristics)"


list(
  icon = "home",
  color = "primary",
  value = paste0(round(comp_df$livingston_RPL_THEME2,2))
)

```

```{r}
#| content: valuebox
#| title: "Theme 3 (Racial/Ethnic Minority Status)"


list(
  icon = "globe",
  color = "secondary",
  value = paste0(round(comp_df$livingston_RPL_THEME3,2))
)

```

```{r}
#| content: valuebox
#| title: "Theme 4 (Transportation and Housing)"


list(
  icon = "car-side",
  color = "warning",
  value = paste0(round(comp_df$livingston_RPL_THEME4,2))
)

```


```

# Plots

```{r}
#| title: Correlation Between SVI Themes




theme_corr <- svi %>% 
  select(SPL_THEME1, SPL_THEME2, SPL_THEME3, SPL_THEME4) %>%
  cor(use = "pairwise.complete.obs")


melted_corr <- melt(theme_corr)


ggplot(melted_corr, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
  labs(title = "Matrix of Correlations Between SVI Themes", x = "", y = "") +
  theme_minimal()
```

```{r}
#| title: Estimated Population of Surrounding Counties

counties <- c('Monroe County', 'Livingston County', 'Wayne County', 'Yates County', 'Genesee County')
popu <- svi %>% 
  filter(COUNTY %in% counties) %>% 
  group_by(COUNTY) %>% 
  summarise(pop = sum(E_TOTPOP)) %>% 
  mutate(COUNTY = fct_reorder(COUNTY, pop))

ggplot(popu,
       mapping = aes(x = COUNTY, y = pop), stat = 'identity')+
  geom_col()
```


