---
title: "Beer Market Data"
author: "Daniel Noone"
date: "12/11/2023"
categories: [beer, drinks, data exploration]
image: 'beer.jfif'

execute: 
  warning: false
  message: false
  
toc: true
---

---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```


# Analyzing the `beer` Data Frame:

```{r}
beer <- read.csv('https://bcdanl.github.io/data/beer_markets.csv')
```

```{r}
rmarkdown::paged_table(beer)
```


## Description of Variables for `beer` Data Frame:

  - `hh`: Identification number of the household;
  
  - `X_purchase_desc`: details of item purchased by hh
  
  - `quantity`: number of items purchased by hh
  
  - `brand`: Brand of beer purchased by hh
      - Bud Light, 
      - Busch Light, 
      - Coors Light, 
      - Miller Lite, or 
      - Natural Light
  
  - `dollar_spent`: Dollar value total of purchase;
  
  - `beer_floz`: volume of beer (fl oz);
  
  - `price_per_floz`: price per fl oz
  
  - `container`: Type of container of beer
  
  - `promo`: Was the item promoted? (coupon, etc...)
  
  - `market`: Scan-track market (or state if rural);
  
  - Other demographic data is present in further variables as well...
  
## Summary Statistics of `beer` Data Frame:


```{r}
skim(beer)
```


# Relationships

## Distribution of `brand`


```{r}
beer1 <- beer %>% 
  group_by(brand) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  mutate(prop = n/sum(n),
         brand_fct = reorder(brand, prop))

ggplot(data = beer1)+
  geom_bar(aes(x = brand_fct, y = prop), stat = 'identity', fill = 'blue')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title = 'Frequency of Beer Brands',
       x = 'Brand',
       y = 'Proportion')
```


#### Comments on analyses...

  - The brands above are in an increasing order of proportion among the brands in `beer` data frame
    - The brands are ordered as follows...
      1. Busch Light
      2. Natural Light
      3. Coors Light
      4. Miller Lite
      5. Bud Light


## `brand` and `dollar spent`


```{r}
beer2 <- beer %>% 
  group_by(brand) %>% 
  summarise(dollar_spent_tot = sum(dollar_spent)) %>% 
  mutate(brand_fct = reorder(brand, dollar_spent_tot)) %>% 
  arrange(-dollar_spent_tot)
View(beer2)

ggplot(beer2)+
  geom_bar(aes(x = brand_fct, y = dollar_spent_tot), stat = 'identity', fill = 'red')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title = 'Dollars Spent Per Brand',
       x = 'Brand',
       y = 'Dollars Spent')
```


#### Comments on analyses....

  - The order is the same as the distribution of `brand` within the `beer` data frame
    - Just now the y-axis is measuring dollars spent by each value of `hh` on each brand
        - Bud Light seems to be most popular all around
        

## `container` and `brand`


```{r}
beer3 <- beer %>% 
  group_by(brand, container) %>% 
  summarise(n = n())
beer3.5 <- beer3 %>% 
  group_by(brand) %>% 
  slice_max(order_by = n, n = 2)
ggplot(data = beer3.5)+
  geom_bar(aes(x = brand, y = n, fill = container), stat = 'identity', position = 'dodge')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title = 'Frequency of Cans and Bottles for Each Brand',
       x = 'Brand',
       y = 'Count')
View(beer3.5)
```


#### Comments on analyses...

  - The bar chart above depicts the frequency of two container types (can or non refillable bottle) per each brand
    - All brands seem to sell more cans than bottles
      - Of them Bud Light sells the most cans
    - Bud light also sells the most bottles

