---
title: "Spotify Listener Data"
author: "Daniel Noone"
date: "12/11/2023"
categories: [music, artists, data exploration]
image: 'spotify.jfif'

execute: 
  warning: false
  message: false
  
toc: true
---

---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```


# Analyzing the `spotify` Data Frame:

```{r}
spotify <- read.csv('https://bcdanl.github.io/data/spotify_all.csv')
```

```{r}
rmarkdown::paged_table(spotify)
```

## Description of Variables for `spotify` Data Frame:

  - `pid`: A unique ID for a specific playlist

  - `playlist_name`: The name of a specific playlist

  - `pos`: Position of the track within a playlist (starting from 0)

  - `artist_name`: Name of the artist on the track

  - `track_name`: name of the track

  - `duration_ms`: duration of the track (in milliseconds)

  - `album_name`: name of the track’s album
  
## Summary Statistics for `spotify`:

```{r}
skim(spotify)
```

# Relationships:

## Top 5 Artists in `artist_name`

```{r}
spot1 <- spotify %>% 
  group_by(artist_name) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  head(5) %>% 
  mutate(prop = n/sum(n),
         art_fct = reorder(artist_name, prop))
View(spot1)

ggplot(data = spot1)+
  geom_bar(aes(x = art_fct, y = prop, fill = artist_name), stat = 'identity')+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=0.5))+
  ggtitle('Frequency of Top 5 Artists')+
  xlab('Artist Name')+
  ylab('Proportion')
length(unique(spotify$artist_name))
```

#### Comments on analyses...

  - Out of the 18866 different artist names in the `spotify` data frame, the top five are...
    1. Drake
    2. Kanye West
    3. Kendrick Lamar
    4. Rihanna
    5. The Weeknd
    
      - These are also in order of greatest to least frequency out of the top 5*

## Top 5 songs in `track_name`

```{r}
spot2 <- spotify %>% 
  group_by(track_name) %>% 
  summarise(n = n()) %>% 
  arrange(-n) %>% 
  head(5) %>% 
  mutate(prop = n/sum(n),
         track_fct = reorder(track_name, prop))

ggplot(data = spot2)+
  geom_bar(aes(x = track_fct, y = prop), stat = 'identity', fill = 'blue')+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=0.5))+
  ggtitle('Top 5 Songs')+
  xlab('Song Name')+
  ylab('Proportion')
length(unique(spotify$track_name))
```


#### Comments on analyses...

  - There are a total of 64115 unique values in `track_name` 
  - Out of those 64115 values, the top 5 are...
    1. Closer
    2. One Dance
    3. HUMBLE
    4. Home
    5. Roses
    
      - In order of decreasing proportion within the top 5 songs*
    
## Most Popular Song in `track_name` for Each of the Top 5 Artists in `artist_name` 


```{r}
spot3 <- spotify %>% 
  filter(artist_name %in% spot1$art_fct) %>% 
  group_by(artist_name, track_name) %>% 
  summarise(n = n())
spot3.5 <- spot3 %>% 
  group_by(artist_name) %>% 
  slice_max(order_by = n, n = 1) %>%  #Selects the top n value for each value of artist, the top n rows of top n values
  ungroup() %>% 
  arrange(-n)
  
ggplot(data = spot3.5) + 
  geom_bar(aes(x = artist_name, y = n, fill = track_name), stat = 'identity')+
  labs(title = 'Most Popular Song For Each Top 5 Artist',
       x = 'Artist',
       y = 'Count')+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```


#### Comments on analyses...

  - Above is the top song`track_name` value and it's count in the `spotify` data frame for each of the top 5 values of `artist_name`
    - Drake : One Dance
    - Kanye West : GOLD DIGGER
    - Kendrick Lamar : HUMBLE
    - Rihanna : Needed Me
    - The Weeknd : Starboy


