---
title: "NYC Restaurant Inspections"
author: "Daniel Noone"
date: "12/10/2023"
categories: [food, data exploration]
image: "nyc.jfif"

execute: 
  warning: false
  message: false
  
toc: true
---

---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Analyzing the `nyc_rest_ins` data frame 🍕

```{r}
nyc_rest_ins <- read.csv('https://bcdanl.github.io/data/DOHMH_NYC_Restaurant_Inspection.csv')
```

```{r}
rmarkdown::paged_table(nyc_rest_ins)
```

## Description of Variables for `nyc_rest_ins` data frame 

  - `CAMIS`: A unique identifier number for each restaurant
  
  - `DBA`: Stands for "doing business as", name of the restaurant
  
  - `BORO`: Indentifies the NYC Borough that the restaurant is located in
    - Manhattan
    - Bronx
    - Brooklyn
    - Queens
    - Staten Island
  
  - `STREET`: Street address of the restaurant
  
  - `CUISINE DESCRIPTION`: Type of cuisine the restaurant sells
  
  - `INSPECTION DATE`: The date the inspection was performed
  
  - `ACTION`: Indicated the action(s) taken as a result of the inspection
  
  - `VIOLATION CODE`: Violation associated with the restaurant's inspection
  
  - `VIOLATION DESCRIPTION`: Descriprion of the violation associated with a restaurant's inspection
  
  - `CRITICAL FLAG`: Indicator of a critical violation as result of inspection
    - Critical
    - Not Critical
    - Not Applicaple
    
  - `SCORE`: Numeric Score for the inspection
  
  - `Grade`: Letter Grade for the Inspection
  
## Summary Statisics for `nyc_rest_ins`:

```{r}
skim(nyc_rest_ins)
```

# Relationships

## `GRADE` and `BORO`

Graphical representation of the proportion of each `GRADE` value in each of the 5 values in `BORO`

```{r}

nyc1 <- nyc_rest_ins %>% 
  group_by(BORO, GRADE) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))

ggplot(data = nyc1,
       mapping = aes(x = GRADE, y = prop))+
  geom_bar(aes(fill = GRADE), stat = 'identity')+
  facet_wrap(.~BORO)
```

#### Comments on analyses...

 - All five boroughs seem to have around equal proportions of each Grade
 
## `CRITICAL FLAG` and `BORO`

Graphical representation of the count of each value of `CRITICAL FLAG` per each value of `BORO`

```{r}
nyc2 <- nyc_rest_ins %>% 
  group_by(BORO, CRITICAL.FLAG) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
View(nyc2)

ggplot(data = nyc2,
       mapping = aes(x = CRITICAL.FLAG, y = n))+
  geom_bar(aes(fill = CRITICAL.FLAG), stat = 'identity')+
  facet_wrap(.~BORO)
```

#### Comments on analyses...

  - The proportion of each value of `CRITICAL.FLAG` within each of the values of `BORO` seems to be about equal
  - This faceted bar chart also gives info about the number of restaurants in each value of `BORO` within this data set. In order of most to least...
    1. Manhattan has the most restaurants
    2. Brooklyn has the second most
    3. Queens the third most
    4. Bronx second to last
    5. Staten Island is last

## `GRADE` and `CRITICAL FLAG`

 Graphical representation of the proportion of each value of `GRADE` per value of `CRITICAL FLAG`

```{r}
nyc3 <- nyc_rest_ins %>% 
  group_by(GRADE, CRITICAL.FLAG, BORO) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
View(nyc3)

ggplot(data = nyc3)+
  geom_bar(mapping = aes(x = CRITICAL.FLAG, y = prop, fill = GRADE), stat = 'identity', position = 'dodge')
```

#### Comments on analyses...

 - The proportion of each value of `GRADE` seems to be almost equal per value of `CRITICAL FLAG` 
 
 - The proportion of grade A seems to be slightly less for restaurants with No Critical Flags
 
 - The proportions of grades B and C appear to be slightly less for restautants with Critical Flags

