---
title: "HW5 - NFL Data"
author: "Daniel Noone"
date: "12/11/2023"
categories: [sports, assignments, data exploration]
image: 'ball.jfif'

execute: 
  warning: false
  message: false
  
toc: true
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

# Link to My GitHub Repository:

My repository can be found [here](https://github.com/dan-noone/dan-noone.github.io)

# Analyzing the `NFL` Data Frame:


```{r}
NFL <- read.csv('https://bcdanl.github.io/data/NFL2022_stuffs.csv')
```

```{r}
rmarkdown::paged_table(NFL)
```


## Description of Variables in `NFL` Data Frame:

  - `play_id`: Play Id number (numeric), when combined with game_id provides id for single play
  
  - `game_id`: 10 digit id for an NFL game
  
  - `drive`: Drive number (numeric)
  
  - `week`: Season week
  
  - `posteam`: abv for team with possession
  
  - `qtr`: Quarter in the game (qtr 5 means overtime)
  
  - `half_seconds_remaining`: seconds remaining in the half
  
  - `down`: Down for a given play
  
  - `pass`: 1 if a pass play, 0 if not
  
  - `wp`: Estimate for winning probability for `posteam` given situation at beginning of the play
  

## Summary Statistics for `NFL` Data Frame


```{r}
skim(NFL)
```


# Questions for Homework 5

## Q2a

  - In data.frame, NFL2022_stuffs, remove observations for which values of posteam is missing.

__Answer__: 

```{r}
NFL <- NFL %>% 
  filter(!(is.na(posteam)))
```


## Q2b

  - Summarize the mean value of pass for each posteam when all the following conditions hold:
    1. `wp` is greater than 20% and less than 75%;
    2. `down` is less than or equal to 2; and
    3. `half_seconds_remaining` is greater than 120.

__Answer__: 

```{r}
NFL_2b <- NFL %>% 
  filter(wp > 0.2 & wp < 0.75 & down <= 2 & half_seconds_remaining > 120)
```


## Q2c


  - Provide both (1) a ggplot code with `geom_point()` using the resulting data.frame in Q2b and (2) a simple comments to describe the mean value of pass for each `posteam`.
    - In the ggplot, reorder the `posteam` categories based on the mean value of `pass` in ascending or in descending order.
    
__Answer__

```{r}
NFL_2c <- NFL_2b %>%
  group_by(posteam) %>% 
  mutate(mean_pass = mean(pass)) %>% 
  ungroup() %>% 
  mutate(posteam_fct = reorder(posteam, mean_pass))

ggplot(data = NFL_2c, mapping = aes(x = mean_pass, y = posteam_fct))+
  geom_point()+
  labs(title = 'Possessing Team and Pass Plays',
       x = 'Percentage of Pass Plays',
       y = 'Team with possession')+
  theme(axis.text.y = element_text(hjust = 1))+
  theme_minimal()
```
On the scatter plot there is a positive relationship between posteam and the percentage of pass plays <br>
Teams such as CIN, KC, LAC. and BUF have a higher occurance of pass plays, <br> 
While teams such as ATL, WAS, CHI, and NO have a lower occurance of pass plays


## Q2d

  - Consider the following data.frame, `NFL2022_epa`:
```{r}
NFL2022_epa <- read_csv('https://bcdanl.github.io/data/NFL2022_epa.csv')
```

  - Create the data.frame, `NFL2022_stuffs_EPA`, that includes

    1. All the variables in the data.frame, `NFL`;
    2. The variables, `passer`, `receiver`, and `epa`, from the data.frame, `NFL2022_epa`. by joining the two data.frames.

  - In the resulting data.frame, `NFL2022_stuffs_EPA`, remove observations with `NA` in `passer`.
  
__Answer__:

```{r}
NFL2022_stuffs_EPA <- NFL %>% 
  left_join(NFL2022_epa)
View(NFL2022_stuffs_EPA)

NFL2022_stuffs_EPA <- NFL2022_stuffs_EPA %>% 
  filter(!(is.na(passer)))
```


## Q2e

  - Provide both (1) a single ggplot and (2) a simple comment to describe the NFL `week`ly trend of weekly mean value of `epa` for each of the following two passers,
    1. `"J.Allen"`
    2. `"P.Mahomes"`
    
__Answer__:

```{r}
NFL_2e <- NFL2022_stuffs_EPA %>% 
  filter(passer == "J.Allen" | passer == "P.Mahomes") %>% 
  group_by(passer, week) %>% 
  summarise(mean_epa = mean(epa))

ggplot(data = NFL_2e,
       mapping = aes(x = week, y = mean_epa, color = passer))+
  geom_line(size = 1, linejoin = 'round')+
  geom_point(size = 1.5)+
  theme(legend.position = 'top')+
  labs(title = "Mean EPA for Two NFL Passers",
       x = 'Week',
       y = 'Mean EPA')+
  scale_color_manual(values = c('blue', 'red'))

```

`P.Mahomes` seems to have a higher mean EPA most of the weeks, while `J.Allen` seems to lag behind


## Q2f

  - Calculate the difference between the mean value of `epa` for `"J.Allen"` the mean value of `epa` for `"P.Mahomes"` for each value of `week`
  
__Answer__:

```{r}
NFL_2f <- NFL_2e %>% 
  spread(passer, mean_epa) %>% 
  mutate(mean_epa_dif = J.Allen - P.Mahomes)

```


## Q2g

  - Summarize the resulting data.frame in `Q2d`, with the following four variables:

    - `posteam`: String abbreviation for the team with possession.
    - `passer`: Name of the player who passed a ball to a receiver by initially taking a three-step drop, and backpedaling into the pocket to make a pass. (Mostly, they are quarterbacks.)
    - `mean_epa`: Mean value of epa in 2022 for each passer
    - `n_pass`: Number of observations for each passer
  - Then find the top 10 NFL passers in 2022 in terms of the mean value of `epa`, conditioning that `n_pass` must be greater than or equal to the third quantile level of `n_pass`.

__Answer__:

```{r}
NFL_2g <- NFL2022_stuffs_EPA %>% 
  group_by(posteam, passer) %>% 
  summarise(mean_epa = mean(epa),
            n_pass = n())

Q3_n_pass <- quantile(NFL_2g$n_pass, 0.75)

NFL_2g_2 <- NFL_2g %>% 
  filter(n_pass >= Q3_n_pass) %>% 
  arrange(-mean_epa) %>% 
  head(10)

```

