---
title: "DANL 310 | Maps with df and Google Maps"
toc: True
author: 'Daniel Noone'
date: '2025-04-16'
categories: ['Notes', 'ggplot', 'Maps']
---

# Using Google Maps API to get GMaps Vizs
```{r}
library(tidyverse)
library(ggmap)
library(gmapsdistance)
library(maps)
library(ggthemes)
library(ggrepel)
google_api <- "AIzaSyA3GEXp7zdSJBAhq8fw30f5z-STVFrZSa0"
register_google(google_api) # for the ggmap package
set.api.key(google_api)     # for the gmapsdistance package
```

```{r}
NYC_Map <- get_map("New York City", 
                   source = "google", 
                   api_key = apiKey, 
                   zoom = 10)

ggmap(NYC_Map) +
  theme_map() 
```

```{r}
ROC_Map <- get_map("Rochester, NY", 
                   source = "google", 
                   api_key = apiKey, 
                   zoom = 9) # the larger value, the more zoom-in

ggmap(ROC_Map) +
  theme_map()
```

```{r}
Geneseo_Map <- get_map("Geneseo, NY",
                       source = "google",
                       api_key = apiKey,
                       zoom = 14) 

locations <- geocode(
  c("South Hall, Geneseo, NY", 
    "Newton Hall, Geneseo, NY")
)

locations <- locations |>
  mutate(label = c("South Hall", "Newton Hall"))

ggmap(Geneseo_Map) + 
  geom_point(data = locations, size = 1, alpha = .5) +
  geom_text_repel(data = locations, 
                  aes(label = label),
                  box.padding = 1.75) +
  theme_map()
```

## Custom fonts with showtext
- Enables google fonts in R plots
```{r}
library(showtext)
font_add_google("Roboto", "roboto")
font_add_google("Lato", "lato")
font_add_google("Poppins", "poppins")
font_add_google("Nunito", "nunito")
font_add_google("Annie Use Your Telescope", "annie")
font_add_google("Pacifico", "pacifico")
showtext_auto()  # Enables Google fonts in R plots.
```
```{r}
ggmap(Geneseo_Map) + 
  geom_point(data = locations, size = 1, alpha = .5) +
  geom_text_repel(data = locations, 
                  aes(label = label),
                  family = "pacifico",
                  box.padding = 1.75) +
  theme_map()
```

## Adding Routes
- `route()`: calculates route between two addresses using GMaps directions
- `mode`: Type of travel (walking, driving, bicycling, transit)
- `geom_path()`: draws route line from route_df coordinates
- `geom_point()`: makes start and end points
```{r}
route_df <- route(
  from = "South Hall, Geneseo, NY",
  to   = "Newton Hall, Geneseo, NY",
  mode = "walking",
  structure = "route"  # returns points along the path 
)

ggmap(Geneseo_Map) +
  geom_path(data = route_df, 
            aes(x = lon, y = lat), 
            color = "#1C4982", size = 2, 
            lineend = "round") +  # round, butt, square
  geom_point(data = locations, 
             aes(x = lon, y = lat), 
             size = 2, color = "darkred") +
  geom_text_repel(data = locations, 
                  aes(x = lon, y = lat, label = label),
                  family = "nunito", 
                  box.padding = 1.75) +
  theme_map()
```

## Different Map Types
```{r}
# Available maptype options:
# "roadmap", "satellite", "terrain", "hybrid"
Geneseo_Map_Hybrid <- get_map("Geneseo, NY",
                              source = "google",
                              api_key = apiKey,
                              zoom = 14, 
                              maptype = "hybrid")

ggmap(Geneseo_Map_Hybrid) +
  geom_path(data = route_df, 
            aes(x = lon, y = lat), 
            color = "#1C4982", 
            size = 2, 
            lineend = "round",  # round, butt, square
            arrow = arrow() # Arrow option
            ) +  
  geom_point(data = locations, 
             aes(x = lon, y = lat), 
             size = 2, color = "darkred") +
  geom_text_repel(data = locations, 
                  aes(x = lon, y = lat, label = label),
                  family = "nunito", 
                  box.padding = 1.75) +
  theme_map()
```

## Multiple Markers w/ Custom Icons
```{r}
library(ggimage)

# Let's suppose we have an icon URL
icon_url <- "https://bcdanl.github.io/lec_figs/marker-icon-red.png"

ggmap(Geneseo_Map) +
  geom_image(data = locations,
             aes(x = lon, y = lat, image = icon_url),
             size = 0.05) +  # tweak size
  geom_text_repel(data = locations, 
                  aes(label = label),
                  box.padding = 1.75) +
  theme_map()
```

# Classwork 9
```{r}
library(tidyverse)
library(skimr)
library(ggthemes)
library(ggmap)
library(sf)
```

## Q1
```{r}
# GeoPackage file (.gkpg), a modern format for map data
nyc_zips_sf <- st_read("https://bcdanl.github.io/data/nyc_zips_sf.gkpg")  # sf format
nyc_dog_bites <- read_csv("https://bcdanl.github.io/data/nyc_dog_bites_all.csv") 
nyc_dog_license <- read_csv(
  'https://bcdanl.github.io/data/nyc_dog_license.csv')
```

```{r}
db <- nyc_dog_bites %>% 
  group_by(zip_code) %>% 
  summarise(n_bites = n()) %>% 
  filter(zip_code %in% nyc_zips_sf$zip_code)

ls <- nyc_dog_license %>% 
  group_by(zip_code) %>% 
  summarise(n_dogs = n()) %>% 
  filter(zip_code %in% nyc_zips_sf$zip_code)

tot <- left_join(ls, db, by = 'zip_code') %>% 
  mutate(n_bites = ifelse(is.na(n_bites), 0, n_bites))

tot <- tot %>% 
  mutate(perc = n_bites/n_dogs)

full <- left_join(nyc_zips_sf, tot, by = 'zip_code')

ggplot()+
  geom_sf(color = 'gray60', data = full, aes(fill = perc))+
  scale_fill_viridis_c(option = 'B')+
  theme_map()+
  theme(legend.position = 'top')
```

## Q2
```{r}
route1 <- route(
  from = "Central Park, NYC",
  to   = "Empire State Building, NYC",
  mode = "driving",
  structure = "route"
)

route2 <- route(
  from = "Empire State Building, NYC",
  to   = "Times Square, NYC",
  mode = "driving",
  structure = "route"
)

route3 <- route(
  from = "Times Square, NYC",
  to   = "Central Park, NYC",
  mode = "driving",
  structure = "route"
)
```
```{r}
NYC_Map <- get_map("Manhattan, NY", 
                   source = "google", 
                   api_key = apiKey, 
                   zoom = 13)

locations <- geocode(
  c("Central Park, NYC", 
    "Empire State Building, NY",
    "Times Square, NYC")
)

locations <- locations |>
  mutate(label = c("Central Park", "Empire State", "Times Square"))

ggmap(NYC_Map) +
  geom_path(data = route1, 
            aes(x = lon, y = lat), 
            color = "red", size = 2, 
            lineend = "round") + 
  geom_path(data = route2, 
            aes(x = lon, y = lat), 
            color = "blue", size = 2, 
            lineend = "round")+
  geom_path(data = route3, 
            aes(x = lon, y = lat), 
            color = "green", size = 2, 
            lineend = "round")+
geom_point(data = locations, 
             aes(x = lon, y = lat), 
             size = 2, color = "darkred") +
  geom_text_repel(data = locations, 
                  aes(x = lon, y = lat, label = label),
                  family = "nunito", 
                  box.padding = 1.75, color = 'navy') +
  theme_map()
```

